---
title: "Data Structures"
author: "Sean Kross"
date: "January 4, 2016"
output: html_document
---

R has four basic data structures: vectors, lists, matricies, and data frames.

### Vectors

Vectors contain data of a single type in 1 dimension. You can create vectors
by using the `c()` function.

```{r}
c(1, 2, 3)
```

Since vectors have types, you can use them with operators to get predictable
results.

```{r}
1 + c(1, 2, 3)
c(1, 2, 3) + c(4, 5, 6)
c(1, 2, 3) * c(4, 5, 6)
TRUE & c(TRUE, FALSE)
TRUE | c(TRUE, FALSE)
c(TRUE, FALSE) & c(TRUE, FALSE)
c(FALSE, TRUE) | c(TRUE, FALSE)
```

As you can see the `&` and `|` operators work with vectors the way `&&` and `||`
work with individual values.

You can extract individual values from vectors by "indexing" them using single
brackets (`[]`).

```{r}
a_vector <- c(6, 7, 8)
a_vector[1]
a_vector[2]
a_vector[3]
```

### Quickly creating vectors

To quickly create a sequence of integers, use the `:` operator:

```{r}
1:7
-4:2
```

To create a vector that repeats a value several times use the `rep()` function:

```{r}
rep(5, times = 4)
rep(3:4, times = 3)
rep(3:4, times = 3, each = 2)
rep(TRUE, times = 3)
rep("four times!", times = 4)
```

To create a sequence of numbers use the `seq()` function:

```{r}
seq(1, 5)
seq(1, 2, by = 0.1)
seq(1, 2, length.out = 5)
```

### More about vectors

You can index vectors by using another vector to specify which values you want
to extract. For example:

```{r}
hundred <- 100:110
hundred[c(1, 3, 4)]
hundred[5:7]
```

You can also combine vectors by using `c()`:

```{r}
onetwothree <- c(1, 2, 3)
fourfivesix <- c(4, 5, 6)
c(onetwothree, fourfivesix)
```

It's also possible to change a value in a vector be re-assigning it:

```{r}
onetwothree
onetwothree[2] <- 100
onetwothree
```

### Matricies

A matrix is like a two dimensional vector. Matricies contain data of a single
type. You can create a matrix by using the `matrix()` function.

```{r}
matrix(1:9, nrow = 3)
matrix(1:9, nrow = 3, byrow = TRUE)
```

Operators work in a predictable way:

```{r}
mat <- matrix(2:5, nrow = 2)
1 + mat
2 * mat
```

Notice that multiplying two matricies does not perform matrix multiplication
according to matrix algebra:

```{r}
mat * mat
```

To multiply matricies using linear algebra use the `%*%` operator:

```{r}
mat %*% mat
```

You can index a matrix with the syntax `a_matrix[row,column]`.

```{r}
big_mat <- matrix(1:9, nrow = 3)
big_mat[1, 2]
```

You can also get a row or a column of a matrix as a vector by leaving out the
`row` or `column`:

```{r}
big_mat[1,]
big_mat[,2]
```

Indexing with vectors also works:

```{r}
big_mat[1:2, 1:2]
big_mat[3, 2:3]
big_mat[c(1, 3), c(1, 3)] # The four "corners" of the matrix
```

You can also re-assign values in a matrix:

```{r}
big_mat
big_mat[2, 2] <- 47
big_mat
```

You can even re-assign an entire row or column:

```{r}
big_mat
big_mat[, 2] <- c(12, 12, 12)
big_mat[1,] <- c(39, 39, 39)
big_mat
```

### Lists

Lists are probably the most universal data structure in R. In its simplest form
a list is a one dimensional data structure that can contain multiple types:

```{r}
list1 <- list(1, TRUE, "three")
list1
```

Lists can be indexed with double brackets:

```{r}
list1[[2]]
list1[[3]]
```

It's also possible to name each element in a list:

```{r}
list2 <- list(first = 85, second = "a word")
list2
```

You can access named list elements with the syntax `a_list[["name"]]` or
`a_list$name`:

```{r}
list2[["first"]]
list2$second
```

You can add named elements to existing lists with a similar synatx:

```{r}
list2[["third"]] <- FALSE
list2$fourth <- "four"
list2
```

