---
title: "`dplyr` & `magrittr`"
author: "Sean Kross"
date: "January 19, 2016"
output: html_document
---

The `dplyr` package can be used for manipulating data frames. You can insatll
it by entering `install.packages("dplyr")` into your R console. After 
installation use `library(dplyr)` to load the package.

```{r, echo=FALSE}
library(dplyr)
library(knitr)
library(magrittr)
```

Retrieve the Lego set data since you're familiar with it:

```{r}
legos <- read.csv("https://raw.githubusercontent.com/seankross/lego/master/data-tidy/legosets.csv", stringsAsFactors = FALSE)
```

### The Basics

You can cause the `legos` data frame to "behave" a little more nicely by
wrapping it with the `tbl_df()` function. I'll do this quickly:

```{r}
legos <- tbl_df(legos)
#kable(head(legos))
```

Let's use `dplyr`'s functions to rearrange the `legos` data frame. You can
re-order the rows of a data frame using the `arrange()` function:

```{r}
# Order `legos` by the number of pieces
#arrange(legos, Pieces)
```

```{r}
legos_simple <- select(legos, Item_Number, Pieces, USD_MSRP)
legos_simple <- filter(legos_simple, !is.na(USD_MSRP))
legos_simple <- mutate(legos_simple, PPP = USD_MSRP / Pieces)
legos_simple <- arrange(legos_simple, desc(PPP))
legos_simple <- filter(legos_simple, PPP < .75)
plot(jitter(rep(0, length(legos_simple$PPP))), legos_simple$PPP,
     xlim = c(-.05, .05))
```

```{r}
legos_simple %<>%
  select(Item_Number, Pieces, USD_MSRP) %>%
  filter(!is.na(USD_MSRP)) %>%
  filter(!is.na(Pieces)) %>%
  mutate(PPP = USD_MSRP / Pieces) %>%
  arrange(desc(PPP)) %>%
  
  tbl_df() %T>%
  print() %>%
  filter(PPP < .75) %>%
  print()
```
#essential
arrange
desc
distinct
filter
mutate
group_by
joins
n
rename
select
slice
summarize
bind_rows
bind_cols
```

