---
title: "Flow Control"
author: "Sean Kross"
date: "January 5, 2016"
output: html_document
---

Okay, now the fun begins! Flow control puts many of the concepts we learned in
[Types & Logic](types.html) to use. We're going to be heavily using the 
`print()` function which prints a string to the R console. Let's test it out
now:

```{r}
print("Hello World!")
```

## What if?

An if statement tests whether an expression is true and *if* the tested
expression is true then a defined series of expressions is executed. A few basic
if statements are illustrated below:

```{r}
if(TRUE){
  print("The expression in the parentheses evaluates to TRUE!")
}

if(FALSE){
  print("The expression in the parentheses evaluates to TRUE!")
}

if(3 < 4){
  print("The expression in the parentheses evaluates to TRUE!")
}

if(3 > 4){
  print("The expression in the parentheses evaluates to TRUE!")
}
```

As you can see the code within the if statement is only executed if the
expression inside of the parentheses evaluates to `TRUE`.

## What else?

You can follow an if statement with an *else* statement which contains a series
of expressions that will only be executed if the expression in parentheses
evaluates to `FALSE`. Here are a few examples:

```{r}
if(TRUE){
  print("The expression in the parentheses evaluates to TRUE!")
} else {
  print("The expression in the parentheses evaluates to FALSE!")
}

if(FALSE){
  print("The expression in the parentheses evaluates to TRUE!")
} else {
  print("The expression in the parentheses evaluates to FALSE!")
}

if(3 < 4){
  print("The expression in the parentheses evaluates to TRUE!")
} else {
  print("The expression in the parentheses evaluates to FALSE!")
}

if(3 > 4){
  print("The expression in the parentheses evaluates to TRUE!")
} else {
  print("The expression in the parentheses evaluates to FALSE!")
}
```

## Chaining if and else

You can chain if and else statements to check multiple logical expressions. You
can see the syntax below:

```{r}
x <- 3
if(x == 2){
  print("x is equal to 2.")
} else if(x == 3){
  print("x is equal to 3.")
} else if(x == 4){
  print("x is equal to 4.")
} else {
  print("x is not equal to 2, 3, or 4.")
}

x <- 5
if(x == 2){
  print("x is equal to 2.")
} else if(x == 3){
  print("x is equal to 3.")
} else if(x == 4){
  print("x is equal to 4.")
} else {
  print("x is not equal to 2, 3, or 4.")
}
```

Once an expression that evaluates to `TRUE` is reached no other expressions in
the chain are evaluated even if later expressions are true. For example:

```{r}
x <- 3
if(x == 2){
  print("x is equal to 2.")
} else if(x == 3){
  print("x is equal to 3.")
} else if(x > 2){ # Notice that this expression is true but is not evaluated.
  print("x is greater than 2.")
} else {
  print("x is not equal to 2, 3, or 4.")
}
```

---

[Home](index.html)