---
title: "Functions"
author: "Sean Kross"
date: "January 5, 2016"
output: html_document
---

In past pages of notes we've seen and used many functions without seriously
defining what functions are. A function is very similar to a mathematical
function. A function may have some input and when a function is
evaluated it either returns a value or it performs some task. In R you can
define your own functions. Below I'll define a very simple function that
returns the number 3:

```{r}
i_return_three <- function(){
  3
}

i_return_three()
```

The last expression of the function is evaluated and the resulting value is
returned. The function `i_return_three()` takes no arguments. Below I'll
illustrate a function that takes one argument:

```{r}
add_one <- function(x){
  x + 1
}

add_one(4)
```

Arguments are variables that are defined within the context of the function.
Variables that are defined outside of the function are ignored if they share the
same name as function arguments. Here's an example of what I'm taking about:

```{r}
y <- 100
arg_y <- function(y) {
  y + 1
}

arg_y(4)

x <- 200
arg_x <- function(z){
  x + 1
}

arg_x(30)
```

You can define as many arguments for a function as you want:

```{r}
add_x_y_z <- function(x, y, z){
  x + y + z
}

add_x_y_z(3, 6, 9)
```

You can also set default values for function arguments:

```{r}
print_a_then_b <- function(a = "a", b = "b"){
  print(a)
  print(b)
}

print_a_then_b()
print_a_then_b("Hello", "World!")
print_a_then_b("World!", "Hello") # it's me...
```

If you explicity specify the value of arguments then the order of arguments
doesn't matter:

```{r}
print_a_then_b(b = "This goes on the bottom", a = "This goes on top")
```

---

### Programming Challange

If you've read through all of the notes for Lecture 1 then you should be able to
complete these challanges:

1. Write a function that prints the integers from 1 to 100.
2. Write a function that prints the integers from 1 to 100, except if the
integer is divisible by three then "Fizz" is printed instead of the number, if
the integer is divisible by five then "Buzz" is printed instead of the number,
and if the integer is divisible by three and five then "FizzBuzz" is printed
instead of the number. You will need to recall how the modulus operator works
(`%%`).
3. Write a function with all of the functionality of #2, but the function also
has two arguments where the first argument is the number where printing should
start and the second argument is the number where printing should end.

---

### [Home](index.html)