---
title: "Lego example"
author: "Sean Kross"
date: "January 8, 2016"
output: html_document
---

First import the data:

```{r}
legos <- read.csv("https://raw.githubusercontent.com/seankross/lego/master/data-tidy/legosets.csv", stringsAsFactors = FALSE)
```

We can look at the top of the data frame using the `head()` function:

```{r}
head(legos)
```

Now let's look at the number of sets that came out each year by using the
`table()` function:

```{r}
lego_table <- table(legos$Year)
lego_table
```

We can turn this table into a data frame pretty easily. The `names()` function
will extract the years from the table, while the `as.vector()` function will
extract the set counts from the table.

```{r}
years <- names(lego_table)
counts <- as.vector(lego_table)
```

Now let's make a new data frame. We'll cast `years` which is a vector of strings
to be a vector of numbers using the `as.numeric()` function:

```{r}
lego_sets_per_year <- data.frame(
  year = as.numeric(years),
  counts = counts
)

head(lego_sets_per_year)
```

Let's make another data frame where we can store the number of sets released
per decade:

```{r}
decades <- seq(1970, 2010, by = 10)
decades

lego_sets_per_decade <- data.frame(
  decade = decades,
  counts = rep(0, times = length(decades))
)
```

Now we can write a for loop that will calculate the number of sets produced
per decade and store that number in the data frame we just created.

```{r}
for(d in decades){
  rows_1970_or_after <- which(lego_sets_per_year >= d)
  rows_1979_or_before <- which(lego_sets_per_year < d + 10)
  
  # See explanation of intersect() below
  rows <- intersect(rows_1970_or_after, rows_1979_or_before)
  
  insert_in_row <- which(lego_sets_per_decade$decade == d)
  lego_sets_per_decade[insert_in_row, 2] <- sum(lego_sets_per_year$counts[rows])
}

lego_sets_per_decade
```

Let's visualize this data frame as a bar plot:

```{r}
barplot(lego_sets_per_decade$counts, names.arg = lego_sets_per_decade$decade)
```

The `intersect()` function takes two vectors as arguments and returns the 
elements that are in both vectors:

```{r}
intersect(c(2, 4, 6, 8, 10), c(1, 2, 4, 8, 16))
```

---

### [Home](index.html)