---
title: "Basic Linear Modeling"
author: "Sean Kross"
date: "January 13, 2016"
output: html_document
---

### Simple Linear Regression

For our exploration of linear modeling we're going to use a dataset that comes
packaged with R called `trees`. This dataset contains information about the
girth, height, and volume of 31 trees. Use `help("trees")` for more infromation
about this dataset. Let's examine the relationship between the girth and the
volume of these trees by plotting them:

```{r}
plot(trees$Girth, trees$Volume, xlab = "Girth", ylab = "Volume", main = "Trees")
```

At first glace it looks like there's a relationship! You could imagine drawing
a straight line through most of the points on this graph, and even the points
that you missed would be close to the line. This "strightness" is called a
linear relationship. R provides functions for modeling linear relationships
and other kinds of relationships between data. The most simple of these
functions is the `lm()` function. To use the `lm()` function you need to provide
a formula, which represents a relationship between variables, and a data frame
that maps the variables in the formula to columns in the data frame. I'll use
`lm()` now to construct a model of the relationship between tree volume and
girth:

```{r}
tree_model1 <- lm(Volume ~ Girth, data = trees)
tree_model1
```

The model itself is an advanced data structure that resembles a list. When I
examine the `tree_model1` variable on the R console it prints the y-axis
intercept of the regression line and the slope of the regression line. Let's
draw the regression line using `abline()`.

```{r}
plot(trees$Girth, trees$Volume, xlab = "Girth", ylab = "Volume", main = "Trees")
abline(tree_model1)
```

```{r}
predictions1 <- predict(tree_model1, data.frame(Girth = 11:14))
predictions1

plot(trees$Girth, trees$Volume, xlab = "Girth", ylab = "Volume", main = "Trees")
abline(tree_model1)
points(11:14, predictions1, col = "red")
```

```{r}
plot(tree_model1$residuals, ylim = c(-10, 10))
abline(h = 0, col = "red")
```

### Multiple Linear Regression

```{r}
plot(trees$Height, trees$Volume, xlab = "Height", ylab = "Volume", main = "Trees")
```

```{r}
tree_model2 <- lm(Volume ~ Girth + Height, data = trees)
tree_model2
```

```{r}
predictions2 <- predict(tree_model2, data.frame(Girth = c(12, 14, 16), 
                                                Height = c(70, 73, 76)))
predictions2
```

```{r}
plot(trees$Girth, trees$Volume, xlab = "Girth", ylab = "Volume", main = "Trees")
points(c(12, 14, 16), predictions2, col = "red")
abline(tree_model1)

plot(trees$Height, trees$Volume, xlab = "Girth", ylab = "Volume", main = "Trees")
points(c(70, 73, 76), predictions2, col = "red")
```

```{r}
plot(tree_model1$residuals, ylim = c(-10, 10))
points(tree_model2$residuals, ylim = c(-10, 10), col = "blue")
abline(h = 0, col = "red")
```

---

### [Home](index.html)