---
title: "Loops"
author: "Sean Kross"
date: "January 5, 2016"
output: html_document
---

Loops allow you to execute a piece of code multiple times. There are two
important kinds of loops in R, for loops and while loops.

## For loops

A for loop iterates a variable through all of the elements of a vector or a
list. Some examples of the syntax are below:

```{r}
for(i in 1:5){
  print(i)
}

number_words <- c("one", "two", "three", "four", "five")

for(i in number_words){
  print(i)
}
```


## While loops

While loops continue to be executed while a specified logical expression
evaluates to `TRUE`. The logical expression in the following example is
`x < 1`.

```{r}
x <- -4
while(x < 1){
  print(x)
  x <- x + 1
}
```

## Controlling loops

The two functions `break` and `next` manipulate the behavior of loops. The
`break` function will immediately exit a loop, while the `next` function will
immediately skip to the next iteration of the loop.

```{r}
for(i in 8:1000000){
  print(i)
  if(i >= 10){
    break
  }
}

for(i in 5:7){
  if(i == 6){
    next
  }
  print(i)
}
```

---

[Home](index.html)