---
title: "Simulation"
author: "Sean Kross"
date: "January 12, 2016"
output: html_document
---

R is wonderful at generating data and conducting simulations of events. In my
opinion there are four main functions for running simulations:

1. `sample()`
2. `replicate()`
3. The `r*()` family of functions.
4. The `*apply()` family of functions.

### `sample()`

If we wanted to simulate flipping a coin or rolling a die, `sample()` is our
go-to function. The `sample()` function in its leanest form can take any vector
as an argument:

```{r, echo=FALSE}
set.seed(1-12-2016)
```

```{r}
sample(1:6)
```

The function randomly picks a number in the vector without replacing it, and
then it continues to randomly pick numbers without replacing them until it
runs out of numbers. To simulate rolling a die, we need to specify the
`replace` argument and the `size` argument.

```{r}
sample(1:6, size = 1, replace = TRUE)
```

This simulates rolling a single die once. To simulate multiple dice rolls we can
increase the `size` argument:

```{r}
sample(1:6, size = 8, replace = TRUE)
```

As you can see the number 4 and 5 are rolled multiple times because we are
sampling elements from the vector with replacement.

### `replicate()`

The `replicate()` function takes two arguments, one is an R expression, and the
other is the number of times you want that R expression to be replicated. For
example if you wanted to compute the mean of 100 rolls of a die you could do
the following:

```{r}
mean(sample(1:6, size = 100, replace = TRUE))
```

Now imagine that you want to compute 100 means of 100 die rolls. You could
accomplish this task using `replicate()`:

```{r}
replicate(100, mean(sample(1:6, size = 100, replace = TRUE)))
```

### The `r*()` functions

`r*()` refers to functions like `rnorm()` and `rbinom()` which generate
random variables from their respective distributions. These functions are
discussed more in-depth in the the [statistical functions](stat.html) notes.
For a list of the distributions that come with R, enter `help("Distributions")`
into the R console.

Below I'll illustrate how you could use `rbinom()` to simulate trials of
flipping a fair coin. The `rbinom()` function needs three arguments: the number
of trials you wish to perform, the number of coins to flip per trial, and the
probability of the coin landing on heads. In the example below I'll simulate
performing 10 trials, each trial flipping 9 coins, and the probability of heads
will be 50%:

```{r}
rbinom(10, 5, .5)
```

As you can see `rbinom()` returns a vector where each element of the vector
represents the number of coins that landed on heads in that trial.

### The `*apply()` functions



---

### [Home](index.html)

