---
title: "Types"
author: "Sean Kross"
date: "December 27, 2015"
output: html_document
---

## What is a type?

Programming languages organize different kinds of data into "types." Programming
languages do this in order to provide structure and logic to programs. There are 
four basic data types in R. 

### Numeric

Numbers in R have the numeric type. Numbers work the way you would expect them
to, in other words you can use them with all of the arithmetic operators:
`+`, `-`, `*`, `/`, and also with operators for exponentiation `^` and
[modulus](https://en.wikipedia.org/wiki/Modulo_operation). Numbers also work
with logical operators which will be explained in the next section.

```{r}
# Addition
3 + 2

# Subtraction
3 - 2

# Multiplication
3 * 2

# Division
3 / 2

# Exponentiation
3^2

# Scientific Notation
10e2

# Modulus
10 %% 3
```

### Logical

Three values in R have the logical type: `TRUE`, `FALSE`, and `NA`. `NA`
represents missing data in R while `TRUE` and `FALSE` represent the value of a
logical expression. Expressions that evaluate to logical values often involve
the logical operators. The following logical operators are usually used to
compare numbers:

- greater than `>`
- greater than or equal `>=`
- less than `<`
- less than or equal `<=`
- exactly equal `==`
- not equal `!=`

Here are some examples using these operators and the resulting logical values:

```{r}
3 > 10
9 >= 8
-3 < 6
4 <= 23
5 >= 5
0.003 == 0.003
28 != 43
```

There are other logical operators that are used for relating logical values:

- and `&&`
- or `||`
- not `!`

Here are some examples using these operators and the resulting logical values:

```{r}
TRUE && TRUE
FALSE && TRUE
TRUE && FALSE
FALSE && FALSE
TRUE || TRUE
FALSE || TRUE
TRUE || FALSE
FALSE || FALSE
!TRUE
!FALSE
```

Logical values can also be used with `==` and `!=` operators.

```{r}
TRUE == TRUE
TRUE == FALSE
TRUE != FALSE
```

All of these logical operators can be combined in expressions. Here is the order
of operations in terms of logical operators, which all come after any mathematical
operations:

1. `>`, `>=`, `<`, `<=`, `==`, `!=`, `!`
2. `&&`
3. `||`

Here are examples:

---

### Exercises:

---

[Home](index.html)