---
title: "Types"
author: "Sean Kross"
date: "December 27, 2015"
output: html_document
---

## What is a type?

Programming languages organize different kinds of data into "types." Programming
languages do this in order to provide structure and logic to programs. There are 
three basic types in R, though other types exist too.

### Numeric

Numbers in R have the numeric type. Numbers work the way you would expect them
to, in other words you can use them with all of the arithmetic operators:
`+`, `-`, `*`, `/`, and also with operators for exponentiation `^` and
[modulus](https://en.wikipedia.org/wiki/Modulo_operation). Numbers also work
with logical operators which will be explained in the next section.

```{r}
# Addition
3 + 2

# Subtraction
3 - 2

# Multiplication
3 * 2

# Division
3 / 2

# Exponentiation
3^2

# Scientific Notation
10e2

# Modulus
10 %% 3
```

### Logical

Three values in R have the logical type: `TRUE`, `FALSE`, and `NA`. `NA`
represents missing data in R while `TRUE` and `FALSE` represent the value of a
logical expression. Expressions that evaluate to logical values often involve
the logical operators. The following logical operators are usually used to
compare numbers:

- greater than `>`
- greater than or equal `>=`
- less than `<`
- less than or equal `<=`
- exactly equal `==`
- not equal `!=`

Here are some examples using these operators and the resulting logical values:

```{r}
3 > 10
9 >= 8
-3 < 6
4 <= 23
5 >= 5
0.003 == 0.003
28 != 43
```

There are other logical operators that are used for relating logical values:

- and `&&`
- or `||`
- not `!`

Here are some examples using these operators and the resulting logical values:

```{r}
TRUE && TRUE
FALSE && TRUE
TRUE && FALSE
FALSE && FALSE
TRUE || TRUE
FALSE || TRUE
TRUE || FALSE
FALSE || FALSE
!TRUE
!FALSE
```

Logical values can also be used with `==` and `!=` operators.

```{r}
TRUE == TRUE
TRUE == FALSE
TRUE != FALSE
```

All of these logical operators can be combined in expressions. Here is the order
of operations in terms of logical operators, which all come after any mathematical
operations:

1. `>`, `>=`, `<`, `<=`, `==`, `!=`, `!`
2. `&&`
3. `||`

Here are examples:

```{r}
3 > 4 && 5 <= 10
4 != 9 || 6 > 0
6 + 9 >= 15 && 8 * 6 > 3
5 + 1 > 6 && 4 != 5 || 9 * 7 < 1000 && TRUE != FALSE
```

#### Exercises

Should the following expressions evaluate to `TRUE` or `FALSE`? Try to work it
out on paper before you check your answer in the R console.

1. `3 > 9 || 8 >= 2^3`
2. `5 > 1 && 3 <= 3 && 4 + 9 * 2 <= 5^3 || (4^2 * 3 > 8 * 5 + 2 || 7^3 > 350) && 9^4 <= 5^6`
3. `[Any R expression that evaluates to a logical value] || TRUE`

### Character

The character type can be thought of as a piece of text, or as a programmer
might call it a string. Strings are text surrounded by single or double quotes.

```{r}
"I am a string!"
'I am also a string!'
```

There's no difference between a string that is single or double quoted as long
as it contains the same text. We can test this with the logical operator for
equality (`==`).

```{r}
"Are we the same?" == 'Are we the same?'
```

The logical operators for equality test whether the contents of the string are
equal. For example:

```{r}
"Are we the same?" == "No way."
"I'm unique!" != "We're different!"
```

### Type Coercion

If you try to use the arithmetic and logical operators we've used with different
types R will attempt to coerce the operator's *arguments* so that the
operator can still work without producing an error. Here are examples of adding
a logical value to a numeric value:

```{r}
0 + TRUE
0 + FALSE
```

As you can see `TRUE` is coerced to 1 and `FALSE` is coerced to 0. This coercion
also takes place when using logical values with arithmetic operators:

```{r}
TRUE + TRUE + FALSE + FALSE
(TRUE + TRUE) * (TRUE + TRUE)
(4 > 3) + (4 < 3)
```

The reverse coercion also takes place where numbers become logical values when
used with logical operators:

```{r}
1 && TRUE
0 && TRUE
```

In this case 0 is always coerced to `FALSE` and any number other than 0 is
coerced to `TRUE`. More examples:

```{r}
45 && 300
0 || 3
0 && 0 || 4 > 7
```

---

### [Home](index.html)