---
title: "Workspace & Files"
author: "Sean Kross"
date: "January 7, 2016"
output: html_document
---

R can interact with files and folders on your computer. R constantly keeps track
of a *working directory* which is a folder in which many R functions automatically
read and write files. Folders and directories are synonymous terms.

You may have never thought about this before, but files and folders on your computer
are structured like a tree. The root of the tree is your hard drive. All files 
and folders are contained within your hard drive, and the series of
folders you might have to navigate to get to a particular file are like the
branches of a tree. This is sequence of directories is often called a *path*.

By default R's inital working directory is your home folder. If you have a Mac
your home folder will have the path `/Users/[Your Username]`, while on Windows
your home folder will have the path `C:\Users\[Your Username]`. Notice that
on Mac the path to your hard drive, also called the root, is `/` and on
Windows it is `C:\`.

The function `getwd()` returns R's current working directory as a string:

```{r, eval=FALSE}
getwd()
```

```
## [1] "/Users/sean"
```

R can create directories and files. You can create a new directory with the
function `dir.create()`:

```{r, eval=FALSE}
dir.create("workspace_and_files")
```

Since our working directory is `/Users/sean`, `dir.create()` creates a new
directory inside of `/Users/sean`. The path of the new directory we created is
`/Users/sean/workspace_and_files`.

Let's change our working directory to `/Users/sean/workspace_and_files`. We can
do this using `setwd()`:

```{r, eval=FALSE}
setwd("workspace_and_files")
```

```{r, eval=FALSE}
getwd()
setwd()
list.files()
dir.create()
dir.exists()
file.create()
file.copy()
file.edit()
file.exists()
file.rename()
file.path()
file.path()
unlink()
```